A Tale of Two Transects

### 2020-02-07 writing/transect_paper.Rmd
variogram table

  - Long amplicons found longer autocorrelation of Bray-Curtis for all fungi, but not ECM.  Is this due to some particular group?
  - Long amplicon also detects some autocorrelation even after 1 year, absent in other tests.
  - Unifrac has less significant autocorrelation.

@@

```{r variofit, results="asis"}
variog_data %>%
  select(mantel_0, mantel_1, variofit_0 = variofit2, variofit_1 = variofitST2, metric, amplicon, tech, guild, algorithm) %>%
  pivot_longer(1:4, names_to = c(".value", "timelag"), names_sep = "_") %>%
  mutate(params = map(variofit, broom::tidy)) %>%
  unnest(params) %>%
  bind_cols(
  pmap_dfr(
    .,
    function(mantel, variofit, term, ...) {
      if (mantel) {
        tryCatch(broom::confint_tidy(object = variofit, parm = term) %>% as.list %>% as.data.frame,
                 error = function(e) tibble(conf.low = NA_real_, conf.high = NA_real_))
      } else {
        tibble(conf.low = NA_real_, conf.high = NA_real_)
      }
    })
  ) %>%
  filter(term %in% c("range", "timerange")) %>%
  mutate(value = ifelse(
    is.na(estimate) | is.na(conf.low) | is.na(conf.high),
    "-",
    sprintf("%.2g (%.2g–%.2g)", estimate, conf.low, conf.high)
  )) %>%
  pivot_wider(names_from = "term", values_from = "value", id_cols = c("guild", "amplicon", "tech", "metric", "algorithm")) %>%
  select(Guild = guild, Amplicon = amplicon, Tech = tech, Algorithm = algorithm, Metric = metric, "Space Range (m)" = range, "Time Range (a)" = timerange) %>%
  arrange(Guild, Amplicon, Tech, Algorithm, Metric) %>%
  kable(booktabs = TRUE, caption = "(ref:variofit)", linesep = c(rep("", 5), "\\addlinespace"))
```

@@

 - Fit looks good for long Pacbio w/Bray-Curtis.
 - Unifrac shows unconvincing distance structure. (No surprise that the model doesn't fit well)
 - Distance structure for short amplicon is really confusing.  Why are the samples 25 m apart so similar? (Even after 1 year?) Could the samples be mixed up? Or is this just a statistical artifact (very few sample pairs have that separation).
 - ECM has shorter range and less anisotropy for all metrics
   - shorter range = more spatially variable
   - less anisotropy = less temporally variable

>>>>

  - Very similar results for different amplicons and technologies
  - Unifrac shows very little distance structure.

@@

```{r variofit, results="asis"}
variog_data %>%
  select(mantel_0, mantel_1, variofit_0 = variofit2, variofit_1 = variofitST2, metric, amplicon, tech, guild, algorithm) %>%
  pivot_longer(1:4, names_to = c(".value", "timelag"), names_sep = "_") %>%
  mutate(params = map(variofit, broom::tidy)) %>%
  unnest(params) %>%
  bind_cols(
    pmap_dfr(
      .,
      function(mantel, variofit, term, ...) {
        if (mantel) {
          tryCatch(broom::confint_tidy(object = variofit, parm = term) %>% as.list %>% inset2("term", term) %>% as.data.frame,
                   error = function(e) tibble(term = term, conf.low = NA_real_, conf.high = NA_real_))
        } else {
          tibble(term = term, conf.low = NA_real_, conf.high = NA_real_)
        }
      }) %>%
      select(-term)
  ) %>%
  filter(term %in% c("range", "timerange")) %>%
  mutate(value = ifelse(
    is.na(estimate) | is.na(conf.low) | is.na(conf.high),
    "-",
    sprintf("%.2g (%.2g–%.2g)", estimate, conf.low, conf.high)
  )) %>%
  filter(timelag == "0" | term == "timerange") %>%
  pivot_wider(names_from = "term", values_from = "value", id_cols = c("guild", "amplicon", "tech", "metric", "algorithm")) %>%
  select(Guild = guild, Amplicon = amplicon, Tech = tech, Algorithm = algorithm, Metric = metric, "Space Range (m)" = range, "Time Range (a)" = timerange) %>%
  arrange(Guild, Amplicon, Tech, Algorithm, Metric) %>%
  kable(booktabs = TRUE, caption = "(ref:variofit)", linesep = c(rep("", 5), "\\addlinespace"))
```

@@

 - Spatial ranges are all consistent, given the wide confidence intervals.
 - For most models the temporal aspect showed poor convergence.
 - Unifrac shows unconvincing distance structure. (No surprise that the model doesn't fit well)
 - similar results for whole community and ECM -- ECM dominate community.
